A comparison of design-based and model-based approaches for finite population spatial data.

### 2021-10-13 inst/manuscript/manuscript.Rmd
section 3 and 4 edits

Treating the data as fixed and incorporating randomness through the sampling design yields estimators having very few other assumptions. Confidence intervals for these types of estimators are typically derived using limiting arguments that incorporate all possible randomizations of sampling units selected via the sampling design. Means and totals, for example, are asymptotically Gaussian by the Central Limit Theorem (under some assumptions). If we repeatedly sample the surface, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true, fixed mean. @sarndal2003model and @lohr2009sampling provide thorough reviews of the design-based approach.

@@

We used a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios. All scenarios used exponential correlation with an effective range of $\sqrt{2}$ for $N = 900$ response values simulated on the unit square in either random locations (Site Locations = Random) or gridded locations (Site Locations = Gridded). The mean for the spatial process generating the response was set to zero.

@@

The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey]. FPBK can be readily performed in `R` with the `sptotal` package [@higham2020sptotal]. We use `sptotal` for both the simulation analysis and the application, estimating parameters with Restricted Maximum Likelihood (REML).

Figure \ref{fig:figeff} shows the relative efficiency of the four approaches from Table \ref{tab:designanalysis} with "IRS-Design" as the baseline:
\mbox{}
\begin{equation*}
\text{E} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
\end{equation*}

where rMSPE is the root-Mean-Squared-Prediction-Error. When there is no spatial correlation (top row), the four approaches have approximately equal rMSPE, even when the assumptions of the model-based approaches are violated. So, using GRTS or using a spatial model does not result in much loss in efficiency even if the response variable is not spatially correlated. When there is high spatial correlation (bottom row), the GRTS-Model approach tends to perform best, but difference in relative efficiency between GRTS-Model and GRTS-Design is not very big. In the lognormal, high partial sill settings (bottom-right facet), GRTS-Design outperforms IRS-Model by a large margin, suggesting that the design decision (whether to use IRS or GRTS) is more important than the analysis decision (whether to analyze using model assumptions or not).

@@

Plot Note: change colours and think about shape
think about legend going on graph

@@

Figure \ref{fig:figconf} shows the coverage for each of the four approaches. We see that the four approaches have somewhat similar coverages in all settings, with GRTS-Design having slightly lower coverage when the response is normal.

In the normal response settings, where assumptions for the model-based approaches and the design-based approaches are satisfied, all approaches have coverage around the nominal 0.95. Because the intervals are symmetric, normal-based intervals, the coverages are generally lower than the nominal 95%. We see that, because the sampling distribution of the mean is asymptotically normal (reference for this? a spatial version of CLT), the coverages when the sample size is 200 are much closer to the nominal 95% than the coverages when the sample size is 50. In general, the more skewed the distribution of the response, the larger the sample size needed to ensure proper coverage of these normal-based intervals.

Do we want to mention stuff like what's in these next bullet points or no?

- for the model-based approach, the more skewed the population is, the higher the sample size needed to satisfy CLT for predicting a mean. The derivation of the BLUP is entirely moment-based (no distribution assumed) but we still need to assume a distribution to estimate spatial parameters and to generate bounds of a prediction interval.
- many confidence intervals generated for design-based approaches also rely on the CLT and the normal distribution to generate the interval. Again, for highly skewed data with a small sample size, this assumption is violated even though all of the assumptions for generating the estimator are valid.

@@

The Environmental Protection Agency (EPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) in the United States to assess the water quality of various bodies of water. We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and quality in lakes in the contiguous United States, to obtain an interval for mean mercury concentration. Although all lakes in the survey were measured in 2012, there may not always be enough time or money to do so. Therefore, we will explore whether or not we can still obtain an adequately precise estimate for the realized mean mercury concentration if we only take a sample of 100 of the 986 lakes.

```{r, message = FALSE, warning = FALSE, echo = FALSE, results = "hide"}
library(tidyverse)
library(here)
library(SpatialDesignVModel)
data(nla)
nla_df
factor(nla_df$TOTALHG_UNITS) ## all same units

@@

```{r figdata, out.width = "100%", fig.cap = "Population distribution of mercury concentration for 986 lakes in the contiguous United States. Thirty-five lakes were dropped from the analysis because they were missing mercury concentration.", message = FALSE, warning = FALSE}
merc <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
  geom_histogram(colour = "black", fill = "white", bins = 15) +
  labs(x = "HG (ng / g)")
# dist <- ggplot(data = nla_nomiss_both, aes(x = RDis_IX)) +
#   geom_histogram(colour = "black", fill = "white", bins = 15) +
#   labs(x = "Lakeshore Disturbance")
merc_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
  geom_point(aes(colour = total_hg)) +
  scale_colour_viridis_c() +
  labs(x = "", y = "", colour = "hg") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank()) +
  coord_quickmap() 
# dist_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
#   geom_point(aes(colour = RDis_IX)) +
#   scale_colour_viridis_c() +
#   labs(x = "", y = "", colour = "Dist") +
#   theme(axis.text.x = element_blank(),
#         axis.text.y = element_blank(),
#         axis.ticks = element_blank())
library(gridExtra)
# grid.arrange(merc, merc_map, nrow = 2)
merc_map
```

```{r, echo = FALSE, results = "hide"}
mean(nla_nomiss_both$total_hg)
# mean(nla_nomiss_both$RDis_IX)
```

Figure \ref{fig:figdata} shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration. Mercury concentration exhibits some spatial correlation, with high mercury concentrations in lakes in the northeast and north central United States. Because we are considering these lakes to be our entire population, we know that the realized mean mercury concentration is 103.2 ng / g.

@@

```{r apptab}
library(knitr)
res_df_drop <- res_df[ ,-2]
names(res_df_drop) <- c("Approach", "Estimate", "SE", "95% LB", "95% UB")
kable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration. The true mean concentration is 103.2 103.2 ng / g")
```

Table \ref{tab:appliedtab} shows the application of a design-based analysis on an IRS, a model-based analysis on an IRS, a design-based analysis on a GRTS sample, and a model-based analysis on a GRTS sample. We see that, for all four analyses, the true realized mean mercury concentration is within the bounds of the 95% intervals. However, we should not generalize the results of this particular realization to any other data set or even to other potential samples of this data set.

<!-- not sure why this is getting cutoff. temporary fix: -->

\pagebreak

But, we do note a couple of patterns. The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not use the spatial correlation in mercury concentration across the contiguous United States. We also see that, for the samples drawn, the both analyses with the GRTS sampling design have a lower standard error than the analyses with the IRS sampling design. We would expect this to be the case for most samples because mercury concentration exhibits spatial correlation so a spatially balanced sample should usually yield a lower standard error. If it is acceptable to have an interval for mean mercury concentration of about 25 ng / g and if we ignore the other variables that the EPA collects information on in these NLA surveys, then the EPA could consider sampling just 50 lakes to save time and money.

>>>>

Treating the data as fixed and incorporating randomness through the sampling design yields estimators having very few other assumptions. Confidence intervals for these types of estimators are typically derived using limiting arguments that incorporate all possible randomizations of sampling units selected via the sampling design. Means and totals, for example, are asymptotically normally distributed (normal) by the Central Limit Theorem (under some assumptions). If we repeatedly sample the surface, then 95% of all 95% confidence intervals constructed from a procedure with appropriate coverage will contain the true, fixed mean. @sarndal2003model and @lohr2009sampling provide thorough reviews of the design-based approach.

@@

We used a crossed design with the simulation parameters given in Table \ref{tab:parmtab} for a total of 36 scenarios. All scenarios used exponential correlation with an effective range of $\sqrt{2}$ for $N = 900$ response values simulated on the unit square in either random locations (Layout = Random) or gridded locations (Layout = Gridded). The mean for the spatial process generating the response was set to zero.

@@

The GRTS algorithm and the local neighborhood variance estimator are available in the \textbf{\textsf{R}} package \texttt{spsurvey} [@dumelle2021spsurvey]. FPBK can be readily performed in `R` with the `sptotal` package [@higham2021sptotal]. We use `sptotal` for both the simulation analysis and the application, estimating parameters with Restricted Maximum Likelihood (REML).

**Mike** For design-based, it's really RMSE -- how should we address this?
Figure \ref{fig:figeff} shows the relative efficiency of the four approaches from Table \ref{tab:designanalysis} with "IRS-Design" as the baseline:
\mbox{}
\begin{equation*}
\text{EFF} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
\end{equation*}

where rMSPE is the root-mean-squared-prediction error. When there is no spatial correlation (top row), the four approaches have approximately equal rMSPE, even when the assumptions of the model-based approaches are violated. So, using GRTS or using a spatial model does not result in much, if any, loss in efficiency even if the response variable is not spatially correlated. When there is high spatial correlation (bottom row), the GRTS-Model approach tends to perform best, but difference in relative efficiency between GRTS-Model and GRTS-Design is small. In the lognormal, high partial sill settings (bottom-right facet), GRTS-Design outperforms IRS-Model by a large margin, suggesting that the design decision (whether to use IRS or GRTS) is more important than the analysis decision (whether to analyze using model assumptions or not).

@@

@@

We also studied 95% interval coverage among the approaches. The design-based 95% confidence intervals and model-based 95% prediction intervals are constructed using the normal distribution. Justification for the design-based intervals lies in the asymptotic normality of totals via the Central Limit Theorem, and justification for the model-based intervals lies in the normality assumption of the errors.  Figure \ref{fig:figconf} shows the 95% interval coverage for each of the four approaches. All four approaches have somewhat similar interval coverage in all settings, with GRTS-Design having slightly lower coverage when the response is normal.

In the normal response settings, all approaches have coverage around 95%. This is expected, as the intervals are also based on the normal distribution In the lognormal response settings, however, all approaches have coverage below 95%. This is also expected, as the intervals are still based on the normal distribution. In the lognormal response settings, interval coverage increases both as the sample size increases and as the strength of spatial dependence increases. This suggests that the larger the sample size and the stronger the spatial dependence, the more resistant these intervals are to departures from normality of the data.

@@

The Environmental Protection Agency (EPA), states, and tribes periodically conduct National Aquatic Research Surveys (NARS) in the United States to assess the water quality of various bodies of water. We will use the 2012 National Lakes Assessment (NLA), which measures various aspects of lake health and quality in lakes in the contiguous United States, to obtain an interval for mean mercury concentration. Although we know the true mean mercury concentration values for the 986 lakes from the 2012 NLA, we will explore whether or not we obtain an adequately precise estimate for the realized mean mercury concentration if we sample only 100 of the 986 lakes.

```{r, message = FALSE, warning = FALSE, echo = FALSE, results = "hide"}
library(tidyverse)
library(here)
library(SpatialDesignVModel)
data(nla_df) # may need to reinstall package
nla_df
factor(nla_df$TOTALHG_UNITS) ## all same units

@@

```{r figdata, out.width = "100%", fig.cap = "Population distribution of mercury concentration for 986 lakes in the contiguous United States.", message = FALSE, warning = FALSE}
merc <- ggplot(data = nla_nomiss_both, aes(x = total_hg)) +
  geom_histogram(colour = "black", fill = "white", bins = 15) +
  labs(x = "HG (ng / g)")
# dist <- ggplot(data = nla_nomiss_both, aes(x = RDis_IX)) +
#   geom_histogram(colour = "black", fill = "white", bins = 15) +
#   labs(x = "Lakeshore Disturbance")
merc_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
  geom_point(aes(colour = total_hg)) +
  scale_colour_viridis_c() +
  labs(x = "", y = "", colour = "hg") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank()) +
  coord_quickmap() +
  theme_bw(base_size = 14)
# dist_map <- ggplot(data = nla_nomiss_both, aes(x = xcoords, y = ycoords)) +
#   geom_point(aes(colour = RDis_IX)) +
#   scale_colour_viridis_c() +
#   labs(x = "", y = "", colour = "Dist") +
#   theme(axis.text.x = element_blank(),
#         axis.text.y = element_blank(),
#         axis.ticks = element_blank())
library(gridExtra)
# grid.arrange(merc, merc_map, nrow = 2)
merc_map
```

```{r, echo = FALSE}
mean_hg <- mean(nla_nomiss_both$total_hg)
# mean(nla_nomiss_both$RDis_IX)
```

Figure \ref{fig:figdata} shows that mercury concentration is right-skewed, with most lakes having a low value of mercury concentration but a few having a much higher concentration. Mercury concentration exhibits some spatial correlation, with high mercury concentrations in lakes in the northeast and north central United States. The realized mean mercury concentration in the 986 lakes is 103.2 ng / g.

@@

```{r apptab, results = "asis"}
res_df_drop <- res_df[ ,-2]
names(res_df_drop) <- c("Approach", "Estimate", "SE", "95% LB", "95% UB")
xtab_app <- xtable(res_df_drop, digits = 1, caption = "\\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration. The true mean concentration is 103.2 ng / g.")
print(xtab_app, include.rownames = FALSE, comment = FALSE)
```

Table \ref{tab:appliedtab} shows the application of a design-based analysis of an IRS sample, a model-based analysis of an IRS sample, a design-based analysis of a GRTS sample, and a model-based analysis of a GRTS sample. For all four analyses, the true realized mean mercury concentration is within the bounds of the 95% intervals. However, we should not generalize the results of this particular realization to any other data set or even to other potential samples of this data set.

<!-- not sure why this is getting cutoff. temporary fix: \pagebreak 
probably because it is an html table -- I used xtable to fix -->

But, we do note a couple of patterns. The design-based IRS analysis shows the largest standard error: a likely reason is that this is the only approach that does not incorporate any spatial information regarding mercury concentration across the contiguous United States. We also see that both approaches using the GRTS sample have a lower standard error than the both approaches using the IRS sample. We would expect this to be the case for most samples because mercury concentration exhibits spatial patterning, so a spatially balanced sample should usually yield a lower standard error.

### 2021-10-13 inst/manuscript/manuscript.tex
section 3 and 4 edits

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A comparison of design-based and model-based approaches for finite population spatial data.},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

Treating the data as fixed and incorporating randomness through the
sampling design yields estimators having very few other assumptions.
Confidence intervals for these types of estimators are typically derived
using limiting arguments that incorporate all possible randomizations of
sampling units selected via the sampling design. Means and totals, for
example, are asymptotically Gaussian by the Central Limit Theorem (under
some assumptions). If we repeatedly sample the surface, then 95\% of all
95\% confidence intervals constructed from a procedure with appropriate
coverage will contain the true, fixed mean. Särndal et al. (2003) and
Lohr (2009) provide thorough reviews of the design-based approach.

@@

We used a crossed design with the simulation parameters given in Table
\ref{tab:parmtab} for a total of 36 scenarios. All scenarios used
exponential correlation with an effective range of \(\sqrt{2}\) for
\(N = 900\) response values simulated on the unit square in either
random locations (Site Locations = Random) or gridded locations (Site
Locations = Gridded). The mean for the spatial process generating the
response was set to zero.

@@

The GRTS algorithm and the local neighborhood variance estimator are
available in the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle
et al., 2021). FPBK can be readily performed in \texttt{R} with the
\texttt{sptotal} package (Higham et al., 2020). We use \texttt{sptotal}
for both the simulation analysis and the application, estimating
parameters with Restricted Maximum Likelihood (REML).

Figure \ref{fig:figeff} shows the relative efficiency of the four
approaches from Table \ref{tab:designanalysis} with ``IRS-Design'' as
the baseline: \mbox{} \begin{equation*}
\text{E} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
\end{equation*}

where rMSPE is the root-Mean-Squared-Prediction-Error. When there is no
spatial correlation (top row), the four approaches have approximately
equal rMSPE, even when the assumptions of the model-based approaches are
violated. So, using GRTS or using a spatial model does not result in
much loss in efficiency even if the response variable is not spatially
correlated. When there is high spatial correlation (bottom row), the
GRTS-Model approach tends to perform best, but difference in relative
efficiency between GRTS-Model and GRTS-Design is not very big. In the
lognormal, high partial sill settings (bottom-right facet), GRTS-Design
outperforms IRS-Model by a large margin, suggesting that the design
decision (whether to use IRS or GRTS) is more important than the
analysis decision (whether to analyze using model assumptions or not).

@@

Plot Note: change colours and think about shape think about legend going
on graph

Figure \ref{fig:figconf} shows the coverage for each of the four
approaches. We see that the four approaches have somewhat similar
coverages in all settings, with GRTS-Design having slightly lower
coverage when the response is normal.

In the normal response settings, where assumptions for the model-based
approaches and the design-based approaches are satisfied, all approaches
have coverage around the nominal 0.95. Because the intervals are
symmetric, normal-based intervals, the coverages are generally lower
than the nominal 95\%. We see that, because the sampling distribution of
the mean is asymptotically normal (reference for this? a spatial version
of CLT), the coverages when the sample size is 200 are much closer to
the nominal 95\% than the coverages when the sample size is 50. In
general, the more skewed the distribution of the response, the larger
the sample size needed to ensure proper coverage of these normal-based
intervals.

Do we want to mention stuff like what's in these next bullet points or
no?

\begin{itemize}
\tightlist
\item
  for the model-based approach, the more skewed the population is, the
  higher the sample size needed to satisfy CLT for predicting a mean.
  The derivation of the BLUP is entirely moment-based (no distribution
  assumed) but we still need to assume a distribution to estimate
  spatial parameters and to generate bounds of a prediction interval.
\item
  many confidence intervals generated for design-based approaches also
  rely on the CLT and the normal distribution to generate the interval.
  Again, for highly skewed data with a small sample size, this
  assumption is violated even though all of the assumptions for
  generating the estimator are valid.
\end{itemize}

@@

The Environmental Protection Agency (EPA), states, and tribes
periodically conduct National Aquatic Research Surveys (NARS) in the
United States to assess the water quality of various bodies of water. We
will use the 2012 National Lakes Assessment (NLA), which measures
various aspects of lake health and quality in lakes in the contiguous
United States, to obtain an interval for mean mercury concentration.
Although all lakes in the survey were measured in 2012, there may not
always be enough time or money to do so. Therefore, we will explore
whether or not we can still obtain an adequately precise estimate for
the realized mean mercury concentration if we only take a sample of 100
of the 986 lakes.

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figdata-1} \caption{Population distribution of mercury concentration for 986 lakes in the contiguous United States. Thirty-five lakes were dropped from the analysis because they were missing mercury concentration.}\label{fig:figdata}
\end{figure}

Figure \ref{fig:figdata} shows that mercury concentration is
right-skewed, with most lakes having a low value of mercury
concentration but a few having a much higher concentration. Mercury
concentration exhibits some spatial correlation, with high mercury
concentrations in lakes in the northeast and north central United
States. Because we are considering these lakes to be our entire
population, we know that the realized mean mercury concentration is
103.2 ng / g.

\begin{longtable}[]{@{}lrrrr@{}}
\caption{\label{tab:appliedtab} Application of design-based and
model-based approaches to the NLA data set on mercury concentration. The
true mean concentration is 103.2 103.2 ng / g}\tabularnewline
\toprule
Approach & Estimate & SE & 95\% LB & 95\% UB\tabularnewline
\midrule
\endfirsthead
\toprule
Approach & Estimate & SE & 95\% LB & 95\% UB\tabularnewline
\midrule
\endhead
IRS-Design & 112.7 & 8.8 & 95.4 & 129.9\tabularnewline
IRS-Model & 110.5 & 7.9 & 95.0 & 125.9\tabularnewline
GRTS-Design & 101.8 & 6.1 & 89.8 & 113.7\tabularnewline
GRTS-Model & 102.3 & 5.9 & 90.8 & 113.9\tabularnewline
\bottomrule
\end{longtable}

Table \ref{tab:appliedtab} shows the application of a design-based
analysis on an IRS, a model-based analysis on an IRS, a design-based
analysis on a GRTS sample, and a model-based analysis on a GRTS sample.
We see that, for all four analyses, the true realized mean mercury
concentration is within the bounds of the 95\% intervals. However, we
should not generalize the results of this particular realization to any
other data set or even to other potential samples of this data set.

\pagebreak

But, we do note a couple of patterns. The design-based IRS analysis
shows the largest standard error: a likely reason is that this is the
only approach that does not use the spatial correlation in mercury
concentration across the contiguous United States. We also see that, for
the samples drawn, the both analyses with the GRTS sampling design have
a lower standard error than the analyses with the IRS sampling design.
We would expect this to be the case for most samples because mercury
concentration exhibits spatial correlation so a spatially balanced
sample should usually yield a lower standard error. If it is acceptable
to have an interval for mean mercury concentration of about 25 ng / g
and if we ignore the other variables that the EPA collects information
on in these NLA surveys, then the EPA could consider sampling just 50
lakes to save time and money.

@@

\leavevmode\hypertarget{ref-higham2020sptotal}{}%
Higham, M., Ver Hoef, J., Bryce, F., 2020. Sptotal: Predicting totals
and weighted sums from spatial data.

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A comparison of design-based and model-based approaches for finite population spatial data.},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

Treating the data as fixed and incorporating randomness through the
sampling design yields estimators having very few other assumptions.
Confidence intervals for these types of estimators are typically derived
using limiting arguments that incorporate all possible randomizations of
sampling units selected via the sampling design. Means and totals, for
example, are asymptotically normally distributed (normal) by the Central
Limit Theorem (under some assumptions). If we repeatedly sample the
surface, then 95\% of all 95\% confidence intervals constructed from a
procedure with appropriate coverage will contain the true, fixed mean.
Särndal et al. (2003) and Lohr (2009) provide thorough reviews of the
design-based approach.

@@

We used a crossed design with the simulation parameters given in Table
\ref{tab:parmtab} for a total of 36 scenarios. All scenarios used
exponential correlation with an effective range of \(\sqrt{2}\) for
\(N = 900\) response values simulated on the unit square in either
random locations (Layout = Random) or gridded locations (Layout =
Gridded). The mean for the spatial process generating the response was
set to zero.

@@

The GRTS algorithm and the local neighborhood variance estimator are
available in the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle
et al., 2021). FPBK can be readily performed in \texttt{R} with the
\texttt{sptotal} package (Higham et al., 2021). We use \texttt{sptotal}
for both the simulation analysis and the application, estimating
parameters with Restricted Maximum Likelihood (REML).

\textbf{Mike} For design-based, it's really RMSE -- how should we
address this? Figure \ref{fig:figeff} shows the relative efficiency of
the four approaches from Table \ref{tab:designanalysis} with
``IRS-Design'' as the baseline: \mbox{} \begin{equation*}
\text{EFF} = \frac{\text{rMSPE of approach}}{\text{rMSPE of IRS-Design}},
\end{equation*}

where rMSPE is the root-mean-squared-prediction error. When there is no
spatial correlation (top row), the four approaches have approximately
equal rMSPE, even when the assumptions of the model-based approaches are
violated. So, using GRTS or using a spatial model does not result in
much, if any, loss in efficiency even if the response variable is not
spatially correlated. When there is high spatial correlation (bottom
row), the GRTS-Model approach tends to perform best, but difference in
relative efficiency between GRTS-Model and GRTS-Design is small. In the
lognormal, high partial sill settings (bottom-right facet), GRTS-Design
outperforms IRS-Model by a large margin, suggesting that the design
decision (whether to use IRS or GRTS) is more important than the
analysis decision (whether to analyze using model assumptions or not).

@@

We also studied 95\% interval coverage among the approaches. The
design-based 95\% confidence intervals and model-based 95\% prediction
intervals are constructed using the normal distribution. Justification
for the design-based intervals lies in the asymptotic normality of
totals via the Central Limit Theorem, and justification for the
model-based intervals lies in the normality assumption of the errors.
Figure \ref{fig:figconf} shows the 95\% interval coverage for each of
the four approaches. All four approaches have somewhat similar interval
coverage in all settings, with GRTS-Design having slightly lower
coverage when the response is normal.

In the normal response settings, all approaches have coverage around
95\%. This is expected, as the intervals are also based on the normal
distribution In the lognormal response settings, however, all approaches
have coverage below 95\%. This is also expected, as the intervals are
still based on the normal distribution. In the lognormal response
settings, interval coverage increases both as the sample size increases
and as the strength of spatial dependence increases. This suggests that
the larger the sample size and the stronger the spatial dependence, the
more resistant these intervals are to departures from normality of the
data.

@@

The Environmental Protection Agency (EPA), states, and tribes
periodically conduct National Aquatic Research Surveys (NARS) in the
United States to assess the water quality of various bodies of water. We
will use the 2012 National Lakes Assessment (NLA), which measures
various aspects of lake health and quality in lakes in the contiguous
United States, to obtain an interval for mean mercury concentration.
Although we know the true mean mercury concentration values for the 986
lakes from the 2012 NLA, we will explore whether or not we obtain an
adequately precise estimate for the realized mean mercury concentration
if we sample only 100 of the 986 lakes.

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/figdata-1} \caption{Population distribution of mercury concentration for 986 lakes in the contiguous United States.}\label{fig:figdata}
\end{figure}

Figure \ref{fig:figdata} shows that mercury concentration is
right-skewed, with most lakes having a low value of mercury
concentration but a few having a much higher concentration. Mercury
concentration exhibits some spatial correlation, with high mercury
concentrations in lakes in the northeast and north central United
States. The realized mean mercury concentration in the 986 lakes is
103.2 ng / g.

\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
Approach & Estimate & SE & 95\% LB & 95\% UB \\ 
  \hline
IRS-Design & 112.7 & 8.8 & 95.4 & 129.9 \\ 
  IRS-Model & 110.5 & 7.9 & 95.0 & 125.9 \\ 
  GRTS-Design & 101.8 & 6.1 & 89.8 & 113.7 \\ 
  GRTS-Model & 102.3 & 5.9 & 90.8 & 113.9 \\ 
   \hline
\end{tabular}
\caption{\label{tab:appliedtab} Application of design-based and model-based approaches to the NLA data set on mercury concentration. The true mean concentration is 103.2 ng / g.} 
\end{table}

Table \ref{tab:appliedtab} shows the application of a design-based
analysis of an IRS sample, a model-based analysis of an IRS sample, a
design-based analysis of a GRTS sample, and a model-based analysis of a
GRTS sample. For all four analyses, the true realized mean mercury
concentration is within the bounds of the 95\% intervals. However, we
should not generalize the results of this particular realization to any
other data set or even to other potential samples of this data set.

But, we do note a couple of patterns. The design-based IRS analysis
shows the largest standard error: a likely reason is that this is the
only approach that does not incorporate any spatial information
regarding mercury concentration across the contiguous United States. We
also see that both approaches using the GRTS sample have a lower
standard error than the both approaches using the IRS sample. We would
expect this to be the case for most samples because mercury
concentration exhibits spatial patterning, so a spatially balanced
sample should usually yield a lower standard error.

@@

\leavevmode\hypertarget{ref-higham2021sptotal}{}%
Higham, M., Ver Hoef, J., Frank, B., Dumelle, M., 2021. Sptotal:
Predicting totals and weighted sums from spatial data.
